*Prochlorococcus marinus* responses to light and oxygen

### 2024-02-23 Code/Manuscript_O2.Rmd
edits to Results & Discussion & Summary

---
title: "*Prochlorococcus marinus* responses to light and oxygen" 
author:
  - name: '**Mireille Savoie**'
    affil: 1
    main: true
    email: msavoie@mta.ca
  - name: Aurora Mattison
    affil: 1, 2
  - name: Julie Nadeau
    affil: 1
  - name: Laurel Genge
    affil: 1
  - name: Sylwia Śliwińska-Wilczewska
    affil: 1
  - name: Maximilian Berthold
    affil: 1, 3
  - name: Naaman Omar
    affil: 1
  - name: Ondřej Prášil
    affil: 1, 4
    main: false
  - name: Amanda M. Cockshutt
    affil: 1, 5
    main: false
  - name: Douglas A. Campbell
    affil: 1
    main: false
    orcid: '0000-0001-8996-5463'
    email: 'dcampbel@mta.ca'
affiliation:
  - num: 1
    address: Department of Biology, Mount Allison University
institute:
  - mta: 'Department of Biology, Mount Allison University, 63B York St., Sackville NB, Canada, E4L 1G7.'
  - num: 2
  - ubc: 'Department of Biology, University of British Columbia'
  - num: 3
  - umwelt: 'Umweltbundesamtm, Dessau-Roßlau, Germany'
  - num: 4
  - alga: 'Algatech, Czech Academy of Science'
  - num: 5
  - sfx: 'Department of Chemistry, St. Frances Xavier University'
date: "`r format(Sys.Date())`"
output:
  bookdown::html_document2:
    code_folding: show
    keep_md: yes
    toc: TRUE
    toc_float: TRUE
    toc_depth: 6
    fig_caption: yes
    pandoc_args: 
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
  bookdown::word_document2:
    code_folding: show
    reference_docx: KnitTemplate.docx
    keep_md: yes
    toc: TRUE
    toc_depth: 6
    fig_caption: yes
    pandoc_args: 
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
always_allow_html: true
bibliography: [Manuscript_O2.bib, RPackages.bib]
csl: plos-one.csl
editor_options: 
  markdown: 
    wrap: 72
---

@@

```{r load libraries, include = FALSE}
# libraries; Note check actual dependencies

library(lubridate)
library(broom)
library(knitr)
library(grid)
library(bookdown)
library(glue)
library(cowplot)
library(ggpubr)
# library(RColorBrewer)
library(tidyverse)
library(png)
library(cowplot)

knitr::write_bib(c("tidyverse", "ggplot2", "Studio", "lubridate","knitr",  "broom", "zoo", "photobiologyWavebands", "RColorBrewer", "data.table","googlesheets4",  "grid", "glue", "bookdown", "photobiology", "mgcv", "minpack.lm", "ggh4x", "cowplot", "egg"), 'RPackages.bib')

@@

## *Prochlorococcus marinus* diversity
*Prochlorococcus marinus*, a genus of Cyanobacteria, is the smallest known photosynthetic prokaryote, with cell diameters ranging from 0.5 to 0.7 µm [@chisholmProchlorococcusMarinusNov1992]. Despite small cell size, *P. marinus* contribute 13 to 48% of net primary production in oligotrophic oceans, corresponding to about 30% of global oxygen production [@partenskyProchlorococcusMarinePhotosynthetic1999].  *P. marinus* growth is currently constrained between latitudes of 40°N to 40°S in open ocean waters, from surface to 300 m depth, thus spanning 3 orders of magnitudes of light [@partenskyProchlorococcusMarinePhotosynthetic1999; @chisholmProchlorococcusMarinusNov1992 ].

@@

@@

XXX Mireille to confirm and summarise XXX

@@

## Naaman GENOMICS METHODS
XXX

@@

XXXMore TextXXX from AuroraXXX
Proteins derived from Clade HLXXX ecotypes of *P. marinus* were detected in OMZ at depths up to 200 meters,  with O~2~ of 15 µM. The extent to which [O~2~] defines
the niches occupied by different *P. marinus* ecotypes, as compared to
potentially covarying environmental variables like photoperiod, light
spectrum, and light level, is poorly described.

@@

@@

The growth rates of *P. marinus* MED4, Clade HLI, increased with higher imposed PAR and longer photoperiods
(Figure \@ref(fig:MED4PAR)), under 250 µM O~2~, across all spectral wavebands. No growth was
observed under any imposed conditions under a 4 h photoperiod.The maximum growth rate
(µ~max~) was
`r PURPARFits %>% filter(Strain == "MED4", WL == "450", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^ achieved under 180 uE blue light and 16 h photoperiod.

Similar to the growth trends under 250 µM O~2~, MED4
maintained at 25 µM O~2~ showed the fastest growth rates when the
photoperiod was 16 h for each spectral waveband, across PAR levels
(Figure \@ref(fig:MED4PAR)).
The µ~max~ was `r PURPARFits %>% filter(Strain == "MED4", WL == "450", O2_uM == "25", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^  achieved under 180 µmol photons
m^-2^ s^-1^ blue light and 16 h photoperiod. The 4 h photoperiod
experiments under white LED light were not performed because no growth
was achieved when grown under an 8 h photoperiod of white LED light.

MED4 did not grow under when sparged to the lowest O~2~
of 2.5 µM (Figure \@ref(fig:MED4PAR)). 2.5 µM O~2~ growth experiments
were not conducted for 4 and 16 h photoperiods, as no reproducible growth occurred
when MED4 was exposed to 8 and 12 h photoperiods.

@@

The GAM model in Figure \@ref(fig:MED4GAM) summarizes MED4 growth responses to red (A,B) or blue (C,D) peak PAR and photoperiod across 2 imposed oxygen concentrations.  Under 250 µM O~2~ MED4 achieved fastest growth rates above peak blue light of ~180 µmol photons m^-2^ s^-1^, and the longest photoperiod of 16, indicated by the 0.64 d^-1^ contour line representing the
90^th^ percentile of maximum achieved growth rate (Figure \@ref(fig:MED4GAM)C).  Growth decreased with decreasing photoperiod and decreasing peak PAR.  Under red light growth was generally slower but the pattern of growth responses to photoperiod and PAR was similar (Figure \@ref(fig:MED4GAM)A).  Note the exclusion of MED4 from growth under 4 h photoperiod under both red and blue light (Figure . Under 25 µM O~2~ MED4 showed similar growth responses, but was excluded from both 4 and 8 h photoperiods.  MED4 did not grow under 25 µM O~2~, so no GAM model was run.

The representative of HLI clade, *P. marinus* MED4, was completely
excluded from growth under the 4 h photoperiod \@ref(fig:MED4GAM)A,C), and showed
increasing growth rates with increasing photoperiod. Thus, MED4 can
potentially exploit longer summer photoperiods in temperate zones, but
will likely remain excluded from growth in winter through its requirement for
more than 4 h of light per day, even if water temperature warms into the Clade HLI tolerance range.

MED4 grew under all PAR levels, so long as the photoperiod was longer than 4 h,
although growth rates were slow under the lowest PAR level of 30 µmol
photons m^-2^ s^-1^. Considering the range of PAR levels that MED4 can
utilize, it is reasonable to suggest that MED4 can inhabit not just
shallow depths, where light levels are high, but also deeper regions
characterized by lower light levels, subject to the limitation of a
photoperiod of at least 4 h even after depth attenuation of light. The photo
regimes of winter temperate zones, due to shorter photoperiods, exclude
MED4 from growth at all depths, however temperate photoperiods and light
levels for the remainder of the year are adequate to support growth.

MED4, grew under all tested spectral wavebands. Growth rates were
generally comparable under blue, white LED and red lights when exposed
to PAR levels greater than 30 µmol photons m^-2^ s^-1^ for each
photoperiod. At the lower PAR of 30 µmol photons m^-2^ s^-1^, MED4 grew
better under blue light. All growth rates for each spectral waveband
increased with increasing photoperiod. The ability to utilize all
spectral wavebands at high PAR levels and longer photoperiods make the
near surface temperate zones available niches for MED4. MED4 ability to
exploit blue light under low PAR levels and variable photoperiods
greater than 4 h allow MED4 to exploit deeper regions in these temperate
zones, where blue light prevails.

@@

The growth rates of *P. marinus* SS120, Clade LLIII, increased with longer photoperiods, under 30 µmol photons m^-2^ s^-1^ peak PAR and 250 µM O~2~ (Figure \@ref(fig:SS120PAR)). No growth was observed under any blue light photoperiods when exposed to peak PAR of 90 µmol photons m^-2^ s^-1^ or greater. Growth rate increased with increasing photoperiods for white and red light under peak PAR of 90 µmol photons m^-2^ s^-1^; however, no growth was observed when the red light photoperiod was 16 h. Experiments under peak PAR
of 180 µmol photons m^-2^ s^-1^ showed no growth under blue nor red lights. Growth rate decreased with longer photoperiods and showed growth inhibition at 12 and 16 h photoperiods under PAR of 180 µmol
photons m^-2^ s^-1^  white LED light.  The µ~max~ was `r PURPARFits %>% filter(Strain == "SS120", WL == "WW", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^  achieved under 90 µmol photons m^-2^ s^-1^ white LED light and 16 h photoperiod.

Similar to the growth trends under the 250 µM O~2~ and PAR of 30 µmol
photons m^-2^ s^-1^ experiments, SS120 maintained at 25 µM O~2~ showed no
growth under a 4 h photoperiod for imposed red spectral
waveband, however under blue light, SS120 was able to grow (Figure
\@ref(fig:SS120PAR)). In contrast to the growth trends of the 250 µM
O~2~ and PAR of 90 µmol photons m^-2^ s^-1^ experiments, SS120 grew
under 4 and 8 h blue light and 16 h red light photoperiods. The growth
rate decreased with increasing photoperiod under PAR of 90 µmol photons
m^-2^ s^-1^ white LED light treatments. Blue light treatments under PAR
of 180 µmol photons m^-2^ s^-1^ showed growth only under an 8 h
photoperiod. The µ~max~ was
`r PURPARFits %>% filter(Strain == "SS120", WL == "450", O2_uM == "25", Photoperiod == 8) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)`
d^-1^ (Table. \@ref(tab:MaxGrowthTable)) achieved under 90 µmol photons
m^-2^ s^-1^ blue light and 8 h photoperiod.

The 25 µM O~2~, less than 16 h photoperiod and 180 µmol photons m^-2^ s^-1^  under white LED light experiments were not performed due to time constraints.

@@

The growth rates of *P. marinus* MIT9313, Clade LLIV increased with longer
photoperiods under low 30 µmol photons m^-2^ s^-1^ peak PAR, under 250 µM O~2~ (Figure
\@ref(fig:MIT9313PAR)).  Under intermediate 90 µmol photons m^-2^ s^-1^ peak PAR growth rates decreased with increasing blue light photoperiod. Blue light did not induce growth at 180 µmol photons m^-2^ s^-1^ peak PAR, while MIT9313 showed only marginal growth under white LED and red light under the 8 h photoperiod . The µ~max~ was `r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ achieved under 30 µmol photons m^-2^ s^-1^ blue light and 16 h photoperiod.

@@

*Prochlorococcus* remain challenging to culture, as their reduced genomes – the smallest of any known oxyphototroph – render them partially dependent upon mutualistic heterotrophic bacteria to  detoxify reactive oxygen species (Moore et al., 2007; Morris et al., 2008). MED4, SS120 and MIT9313 have been successfully cultured in laboratories, and used to show that ecotypic classifications correspond to biochemical differences among strains. CITATONS OUR LAB and OTHERS.

Under full atmospheric [O~2~] and bluelight, LL clades of *Prochlorococcus* are restricted to growth under low light, in part because they suffer photoinhibition of Photosystem II (PSII) through several paths, including direct absorbance of UV or blue light, in parallel with generation of Reactive Oxygen Species (ROS) if the electron flow is slowed, [@aroPhotoinhibitionPhotosystemII1993], producing  damaging singlet oxygen (^1^O~2~) [@aroPhotoinhibitionPhotosystemII1993; @soitamoPhotoinhibitionMarinePicocyanobacteria2017; @murphyPhotoinactivationPhotosystemII2017; @hakalaEvidenceRoleOxygenevolving2005].
Repair of inhibited PSII process relies on the removal of damaged PsbA (Mann et al., 2000; Adam et al., 2005), followed by reassembly with newly synthesized PsbA [@nixonRecentAdvancesUnderstanding2010]. 
Degradation of PsbA is a rate-limiting step in recovery from photoinhibition (Kanervo et al., 1993), mediated largely by a heterohexamer of (FtsH12)3, a membrane-bound (Sacharz et al., 2015)(Zak et al., 2001) metalloprotease  (Chiba et al., 2002) (Yoshioka-Nishimura and Yamamoto, 2014).  (Adam et al., 2005; Boehm et al., 2012) (Nixon et al. (2010)) (Komenda et al., 2007). (Pisareva et al., 2007; Sacharz et al., 2015),  (Mann et al., 2000), (AMANDA REF from ALGATECH).

*Prochlorococcus* genomes encode 4 FtsH proteins, henceforth referred to as FtsH1-4, homologs to the characterized FtsH isoforms of the model cyanobacterium *Synechocystis* FtsH, and with presumably parallel functions (table \@ref(tab:FtsHTable)). Upon a shift to higher light HLIII MED4 upregulated expression of FtsH1 and FtsH2 [@bonisteelStrainSpecificDifferences2018], homologs to the *Synechocystis* slr0228 and slr1604, implicated in PSII repair. In contrast representative LLIV strain MIT9313 showed less overall expression of the FtsH proteases, and thus has fewer FtsH serving each photosystem.  Furthermore MIT9313 expressed primarily FtsH3, homologous to Sll1463 involved in PSI biogenesis, and FtsH expression did not increase in response to light stress. Through adaptation to steady low light, LLIV *Prochlorococcus* instead allocates resources processes other than dynamic regulation of PSII repair.

@@

## **Prochlorococcus* genomic capacities for electron dissipation and DNA repair
NAAMAN FIGURE
Add Clade Code
Make visual grammar consistent.
Drop 'title'
X axis scale annotated in mMol O2, lines in uM, make consistent.

Figure \@ref(fig:ProchlorococcusEnzymeKms) shows genes encoding enzymes for  *P. marinus* strains from Clades HLI, LLII, LLIII and LLII [@omarAnnotationGenesEncoding2023], compared to the measured or inferred K~M~ for [O~2~] for each enzyme CITATIONS. [@bergResponsesPsbAHli2011] previously found that MED4 increases expression of alternative oxidase ('ubiquinol oxidase (non electrogenic)')  to cope with changes in light by dissipating electrons from the inter-system transport chain, while the genome scan shows SS120 and MIT9313 lack this gene, and therefore, lack an oxygen-dependent path to cope with changing excitation. The approximate K~M~ for [O~2~] of ~ 25 µM for ubiquinol oxidase (non electrogenic) is comparable to the lower limit for growth of MED4 in our experiments.  We suggest that dependence upon this enzyme excludes MED4 from low oxygen zones.  Conversely, a gene encoding (S)-2-hydroxy-acid oxidase is encoded in the MIT9313 genome.  (S)-2-hydroxy-acid oxidase catalyzes the reaction of 2-hydroxy acid with O~2~ to produce toxic H~2~O~2~ [@InformationEC15]. (S)-2-hydroxy-acid oxidase has an approximate K~M~ for [O~2~] of ~ 250 µM, and produces H~2~O~2~, so growth at lower [O~2~] may protect MIT9313 from auto-intoxication from production of H~2~O~2~. We hypothesize that under our 250 µM O~2~ and higher blue light, *P. marinus* MIT9313 exhibited photoinhibition, resulting from the inactivation of PSII caused by the production of the reactive oxygen species, hydrogen peroxide. This photoinhibition is further compounded by the limited inducible repair mechanism for PSII, due to the absence of FtsH 2 and 3  expression in *P. marinus* MIT9313 [@bonisteelStrainSpecificDifferences2018].  We hypothesize that under the conditions of our high light and 2.5 µM and 25 µM O~2~ experiments, the activity of the 2-hydroxy-acid oxidase enzyme is suppressed. As a result, the catalyzed production of hydrogen peroxide is inhibited, leading to minimal PSII damage, allowing MIT9313 to avoid photoinhibition and circumvent limitations on PSII repair.  Figure \@ref(fig:ProchlorococcusEnzymeKms) also shows that *P. marinus* SS120 is the only tested ecotype to lack the pyridoxal 5'-phosphate synthase enzyme. The pyridoxal 5'-phosphate synthase enzyme is an important cofactor in the biosynthesis of vitamin B~6~ [@sangIdentificationPyridoxinePyridoxamine2007]. Vitamin B~6~ is a potential antioxidant and can effectively quench singlet oxygen [@bilskiVitaminB6Pyridoxine2000]. The absence of the pyridoxal 5'-phosphate synthase enzyme may explain why *P. marinus* SS120 does not grow as well as *P. marinus* MIT9313 when exposed to high light stress under 25 µM O~2~ and not at all under 2.5 µM O~2~.

@@

# Acknowledgements  {.unnumbered}
Jonah Sheinin assisted with growth of some cultures for protein analyses, while Carlie Barnhill assisted with code for import of multicultivator growth data files.

>>>>

---
title: "*Prochlorococcus marinus* responses to light and oxygen" 
author:
  - name: '**Mireille Savoie**'
    affil: 1
    main: true
    email: msavoie@mta.ca
  - name: Aurora Mattison
    affil: 1, 2
  - name: Julie Nadeau
    affil: 1
  - name: Laurel Genge
    affil: 1
  - name: Sylwia Śliwińska-Wilczewska
    affil: 1
  - name: Maximilian Berthold
    affil: 1, 3
  - name: Naaman Omar
    affil: 1
  - name: Ondřej Prášil
    affil: 1, 4
    main: false
  - name: Amanda M. Cockshutt
    affil: 1, 5
    main: false
  - name: Douglas A. Campbell
    affil: 1
    main: false
    orcid: '0000-0001-8996-5463'
    email: 'dcampbel@mta.ca'
affiliation:
  - num: 1
    address: Department of Biology, Mount Allison University
institute:
  - mta: 'Department of Biology, Mount Allison University, 63B York St., Sackville NB, Canada, E4L 1G7.'
  - num: 2
  - ubc: 'Department of Biology, University of British Columbia'
  - num: 3
  - umwelt: 'Umweltbundesamtm, Dessau-Roßlau, Germany'
  - num: 4
  - alga: 'Algatech, Czech Academy of Science'
  - num: 5
  - sfx: 'Department of Chemistry, St. Frances Xavier University'
date: "`r format(Sys.Date())`"
output:
  bookdown::html_document2:
    code_folding: show
    keep_md: yes
    toc: TRUE
    toc_float: TRUE
    toc_depth: 6
    fig_caption: yes
    pandoc_args: 
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
  bookdown::word_document2:
    code_folding: show
    reference_docx: KnitTemplate.docx
    keep_md: yes
    toc: FALSE
    toc_depth: 6
    fig_caption: yes
    pandoc_args: 
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
always_allow_html: true
bibliography: [Manuscript_O2.bib, RPackages.bib]
csl: plos-one.csl
editor_options: 
  markdown: 
    wrap: 72
---

@@

```{r load libraries, include = FALSE}
# libraries; Note check actual dependencies
# Naaman add packages?
library(knitr)
library(tidyverse)
library(broom)
library(grid)
library(bookdown)
library(glue)
library(cowplot)
library(ggpubr)

library(png)
library(cowplot)

knitr::write_bib(c("knitr", "tidyverse", "lubridate", "ggplot2", "broom", "grid", "bookdown", "glue", "cowplot", "ggpubr", "ggh4x", "png", "Studio", "zoo", "photobiology", "photobiologyWavebands", "data.table", "googlesheets4",  "mgcv", "minpack.lm",  "egg", 'RPackages.bib'))

@@

## *Prochlorococcus marinus* diversity {.unnumbered}
*Prochlorococcus marinus*, a genus of Cyanobacteria, is the smallest known photosynthetic prokaryote, with cell diameters ranging from 0.5 to 0.7 µm [@chisholmProchlorococcusMarinusNov1992]. Despite small cell size, *P. marinus* contribute 13 to 48% of net primary production in oligotrophic oceans, corresponding to about 30% of global oxygen production [@partenskyProchlorococcusMarinePhotosynthetic1999].  *P. marinus* growth is currently constrained between latitudes of 40°N to 40°S in open ocean waters, from surface to 300 m depth, thus spanning 3 orders of magnitudes of light [@partenskyProchlorococcusMarinePhotosynthetic1999; @chisholmProchlorococcusMarinusNov1992 ].

@@

Current niches of *P. marinus* span ocean water columns
[@chisholmNovelFreelivingProchlorophyte1988; @veldhuisVerticalDistributionPigment1990; @partenskyProchlorococcusMarinePhotosynthetic1999] and extend into regions with varying dissolved oxygen concentrations
[@partenskyProchlorococcusAdvantagesLimits2010; @goerickeNovelNicheProchlorococcus2000] and even into anoxic oceans [@lavinNovelLineagesProchlorococcus2010], [@ulloaCyanobacteriumProchlorococcusHas2021; @garcia-robledoCrypticOxygenCycling2017].

@@

@@

## *Prochlorococcus* comparative genomics {.unnumbered}

NAAMAN

@@

XXXMore TextXXX from Amanda & AuroraXXX
Proteins derived from Clade HLXXX ecotypes of *P. marinus* were detected in OMZ at depths up to 200 meters,  with O~2~ of 15 µM. The extent to which [O~2~] defines the niches occupied by different *P. marinus* ecotypes, as compared to potentially covarying environmental variables like photoperiod, light spectrum, and light level, is poorly described.

@@

Guided by the evidence of ocean distributions of proteins from *Prochlorococcus* we set up a matrix of photoperiods, PAR, spectral bands, and [O~2~] to approximate current, and potential future, latitudinal, depth and seasonal niches for *Prochlorococcus* strains.  We earlier found [@murphyPhotoinactivationPhotosystemII2017] that growth under red light could prove mechanistically informative as to factors limiting *Prochlorococcus* growth, so we included that spectral band even though it is not characteristic of representative *Prochlorococcus* niches.  Although *Prochlorococcus* is currently limited to a narrow range of surface photoperiods, potential poleward latitudinal expansions, in combination with attenuation of light with depth, mean *Prochlorococcus* currently, or potentially, encounters a wide range of photoperiods.  Our growth rate determinations generally agreed with those from @mooreComparativePhysiologySynechococcus1995, for white LED and 250 µM O~2~, but our study is, we believe, the first to analyze the interactive growth responses of *Prochlorococcus* strains to varying [O~2~], spectral wavebands and photoperiods.

@@

Under 250 µM O~2~ the growth rates of *P. marinus* MED4, Clade HLI, increased with higher imposed PAR and longer photoperiods
(Figure \@ref(fig:MED4PAR)), across all spectral wavebands. No growth was observed under any imposed conditions under a 4 h photoperiod. The maximum growth rate (µ~max~) was `r PURPARFits %>% filter(Strain == "MED4", WL == "450", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ achieved under 180 uE blue light and 16 h photoperiod.

Similar to growth trends under 250 µM O~2~, MED4 maintained at 25 µM O~2~ showed fastest growth when the
photoperiod was 16 h for each spectral waveband, across PAR levels (Figure \@ref(fig:MED4PAR)).
The µ~max~ was `r PURPARFits %>% filter(Strain == "MED4", WL == "450", O2_uM == "25", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^  achieved under 180 µmol photons m^-2^ s^-1^ blue light and 16 h photoperiod. The 4 h photoperiod
experiments under white LED light were not performed because no growth was achieved when grown under an 8 h photoperiod of white LED light.

MED4 did not grow under when sparged to the lowest [O~2~] of 2.5 µM (Figure \@ref(fig:MED4PAR)). 2.5 µM O~2~ growth experiments
were not conducted for 4 and 16 h photoperiods, as no reproducible growth occurred when MED4 was exposed to 8 and 12 h photoperiods.

@@

The GAM model in Figure \@ref(fig:MED4GAM) summarizes MED4 growth responses to red (A,B) or blue (C,D) peak PAR and photoperiod across 2 imposed oxygen concentrations.  Under 250 µM O~2~ MED4 achieved fastest growth rates above peak blue light of ~180 µmol photons m^-2^ s^-1^, and the longest photoperiod of 16, indicated by the 0.64 d^-1^ contour line representing the 90^th^ percentile of maximum achieved growth rate (Figure \@ref(fig:MED4GAM)C).  Growth decreased with decreasing photoperiod and decreasing peak PAR. Under red light growth was generally slower but the pattern of growth responses to photoperiod and PAR was similar (Figure \@ref(fig:MED4GAM)A).  Note the exclusion of MED4 from growth under 4 h photoperiod under both red and blue light \@ref(fig:MED4GAM). Under 25 µM O~2~ MED4 showed similar growth responses, but was excluded from both 4 and 8 h photoperiods.  MED4 did not grow under 25 µM O~2~, so no GAM model was run. 
Considering the range of PAR levels, and spectral bands that MED4 can utilize, MED4 can inhabit not just shallow depths, where light levels are high, but also deeper regions, characterized by lower level of blue light, subject to the limitation of a photoperiod of more than 4 h, even after depth attenuation of light. The photoregimes of winter temperate zones, due to shorter photoperiods, exclude MED4 from growth at any depth, however temperate photoperiods and light levels for the remainder of the year are potentially adequate to support MED4 growth, if water temperatures warm into the Clade HLI tolerance range.

@@

Under 250 µM O~2~ growth rates of *P. marinus* SS120, Clade LLIII, increased with longer photoperiods, under 30 µmol photons m^-2^ s^-1^ peak PAR and (Figure \@ref(fig:SS120PAR)). No growth was observed under any blue light photoperiods when exposed to peak PAR of 90 µmol photons m^-2^ s^-1^ or greater. Growth rate increased with increasing photoperiods for white and red light under peak PAR of 90 µmol photons m^-2^ s^-1^.  Growth rate decreased with longer photoperiods and showed growth inhibition at 12 and 16 h photoperiods under PAR of 180 µmol photons m^-2^ s^-1^  white, red or blue LED light.  The µ~max~ was `r PURPARFits %>% filter(Strain == "SS120", WL == "WW", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^  achieved under 90 µmol photons m^-2^ s^-1^ white LED light and 16 h photoperiod.

Under 25 µM O~2~ and PAR of 30 µmol photons m^-2^ s^-1^growth trends were similar to 250 µM O~2~. SS120 maintaine showed no
growth under a 4 h photoperiod for red spectral waveband, however under blue light, SS120 was able to grow under a 4 h photoperiod (Figure
\@ref(fig:SS120PAR)). In contrast to the growth trends of the 250 µM O~2~ and PAR of 90 µmol photons m^-2^ s^-1^ experiments, SS120 grew under 4 and 8 h blue light and 16 h red light photoperiods. The growth rate decreased with increasing photoperiod under PAR of 90 µmol photons
m^-2^ s^-1^ white LED light treatments. Blue light treatments under PAR of 180 µmol photons m^-2^ s^-1^ showed growth only under an 8 h
photoperiod. The µ~max~ was `r PURPARFits %>% filter(Strain == "SS120", WL == "450", O2_uM == "25", Photoperiod == 8) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ (Table. \@ref(tab:MaxGrowthTable)) achieved under 90 µmol photons
m^-2^ s^-1^ blue light and 8 h photoperiod. The 25 µM O~2~, less than 16 h photoperiod and 180 µmol photons m^-2^ s^-1^  under white LED light experiments were not performed due to time constraints.

@@

The growth rates of *P. marinus* MIT9313, Clade LLIV increased with longer
photoperiods under low 30 µmol photons m^-2^ s^-1^ peak PAR, under 250 µM O~2~ (Figure
\@ref(fig:MIT9313PAR)).  Under intermediate 90 µmol photons m^-2^ s^-1^ peak PAR growth rates decreased with increasing blue light photoperiod. Blue light did not induce growth at 180 µmol photons m^-2^ s^-1^ peak PAR, while MIT9313 showed only marginal growth under white LED and red light under the 8 h photoperiod, consistent with [@moorePhotophysiologyMarineCyanobacterium1999]. The µ~max~ was `r PURPARFits %>% filter(Strain == "MIT9313", WL == "450", O2_uM == "250", Photoperiod == 16) %>% select(ReplicateMean_deltaOD_Lmu_corr_day) %>% max() %>% round(digits = 2)` d^-1^ achieved under 30 µmol photons m^-2^ s^-1^ blue light and 16 h photoperiod.

@@

*Prochlorococcus* remain challenging to culture, as their reduced genomes – the smallest of any known oxyphototroph – render them partially dependent upon mutualistic heterotrophic bacteria to  detoxify reactive oxygen species [@morrisFacilitationRobustGrowth2008; @morrisDependenceCyanobacteriumProchlorococcus2011]. MED4, SS120 and MIT9313 have been successfully cultured in laboratories  [@mooreComparativePhysiologySynechococcus1995; @mooreCulturingMarineCyanobacterium2007], and used to show that ecotypic classifications correspond to biochemical differences among strains [@bonisteelStrainSpecificDifferences2018]. CITATONS OTHERS.

Under full atmospheric [O~2~] and bluelight, LL clades of *Prochlorococcus* are restricted to growth under low light, in part because they suffer photoinhibition of Photosystem II (PSII) through several paths, including direct absorbance of UV or blue light, in parallel with generation of Reactive Oxygen Species (ROS) if the electron flow is slowed, [@aroPhotoinhibitionPhotosystemII1993], producing damaging singlet oxygen (^1^O~2~) [@aroPhotoinhibitionPhotosystemII1993; @soitamoPhotoinhibitionMarinePicocyanobacteria2017; @murphyPhotoinactivationPhotosystemII2017; @hakalaEvidenceRoleOxygenevolving2005]. Repair of photoinactivated PSII relies on the removal of damaged PsbA (Mann et al., 2000; Adam et al., 2005), followed by reassembly with newly synthesized PsbA [@nixonRecentAdvancesUnderstanding2010]. Degradation of PsbA is a rate-limiting step in recovery from photoinhibition (Kanervo et al., 1993), mediated largely by a heterohexamer of (FtsH12)3, a membrane-bound (Sacharz et al., 2015)(Zak et al., 2001) metalloprotease  (Chiba et al., 2002) (Yoshioka-Nishimura and Yamamoto, 2014).  (Adam et al., 2005; Boehm et al., 2012) (Nixon et al. (2010)) (Komenda et al., 2007). (Pisareva et al., 2007; Sacharz et al., 2015),  (Mann et al., 2000), (AMANDA REF from ALGATECH).

*Prochlorococcus* genomes encode 4 FtsH proteins, henceforth referred to as FtsH1-4, homologs to the characterized FtsH isoforms of the model cyanobacterium *Synechocystis* FtsH, and with presumably parallel functions (table \@ref(tab:FtsHTable)). Upon a shift to higher light HLIII MED4 upregulated expression of FtsH1 and FtsH2 [@bonisteelStrainSpecificDifferences2018], homologs to the *Synechocystis* slr0228 and slr1604, implicated in PSII repair. In contrast representative LLIV strain MIT9313 showed less overall expression of the FtsH proteases, and thus has fewer FtsH serving each photosystem.  Furthermore, MIT9313 expressed primarily FtsH3, homologous to *Synechocystis* slr1463 involved in PSI biogenesis, and  FtsH expression did not increase in response to light stress in MIT9313. Through adaptation to steady low light, Clade LLIV *Prochlorococcus* instead allocate resources to processes other than dynamic regulation of PSII repair.

@@

## **Prochlorococcus* genomic capacities for electron dissipation and DNA repair

Figure \@ref(fig:ProchlorococcusEnzymeKms) shows genes encoding enzymes for  *P. marinus* strains from Clades HLI, LLII, LLIII and LLII [@omarAnnotationGenesEncoding2023], compared to the measured or inferred K~M~ for [O~2~] for each enzyme CITATIONS. MED4 increases expression of alternative oxidase ('ubiquinol oxidase (non electrogenic)')  to cope with changes in light [@bergResponsesPsbAHli2011], by dissipating electrons from the inter-system transport chain. The approximate K~M~ for [O~2~] of ~ 25 µM for ubiquinol oxidase (non electrogenic) is comparable to the lower limit for growth of MED4 in our experiments.  We suggest that dependence upon this enzyme excludes MED4 from low oxygen zones.  The genome scan shows SS120 and MIT9313 lack this gene, and therefore, lack this oxygen-dependent path to cope with changing excitation.  Conversely, a gene encoding (S)-2-hydroxy-acid oxidase is encoded in the MIT9313 genome.  (S)-2-hydroxy-acid oxidase catalyzes the reaction of 2-hydroxy acid with O~2~ to produce toxic H~2~O~2~ [@InformationEC15]. (S)-2-hydroxy-acid oxidase has an approximate K~M~ for [O~2~] of ~ 250 µM, and produces H~2~O~2~, so growth at lower [O~2~] may protect MIT9313 from auto-intoxication from production of H~2~O~2~. We hypothesize that under 250 µM O~2~ and higher blue light, *P. marinus* MIT9313 suffered photoinhibition, resulting from the inactivation of PSII caused by the production of the reactive oxygen species, hydrogen peroxide. This photoinhibition is compounded by the limited inducible repair mechanism for PSII, due to the absence of FtsH 2 and 3  expression in *P. marinus* MIT9313 [@bonisteelStrainSpecificDifferences2018].  We hypothesize that under the conditions of our high light and 2.5 µM and 25 µM O~2~ experiments, the activity of the (S)-2-hydroxy-acid oxidase enzyme is suppressed. As a result, the catalyzed production of hydrogen peroxide is inhibited, leading to less PSII damage, allowing MIT9313 to avoid photoinhibition and circumvent its limitations on PSII repair.  Figure \@ref(fig:ProchlorococcusEnzymeKms) also shows that *P. marinus* SS120 is the only tested ecotype to lack the pyridoxal 5'-phosphate synthase enzyme. The pyridoxal 5'-phosphate synthase enzyme is an important cofactor in the biosynthesis of vitamin B~6~ [@sangIdentificationPyridoxinePyridoxamine2007]. Vitamin B~6~ is a potential antioxidant and can effectively quench singlet oxygen [@bilskiVitaminB6Pyridoxine2000]. The absence of the pyridoxal 5'-phosphate synthase enzyme may explain why *P. marinus* SS120 does not grow as well as *P. marinus* MIT9313, when exposed to high light stress under 25 µM O~2~ and not at all under 2.5 µM O~2~.

@@

# Acknowledgements  {.unnumbered}
Jonah Sheinin (Mount Allison Student) assisted with growth of some cultures for protein analyses, while Carlie Barnhill (Mount Allison Student) assisted with code for import of multicultivator growth data files.
